\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-15 paper.qmd
Add logic to remove long edges and add error calculation

@@

#(n-neighbors: 50)
plot_list1_umap <- plot_UMAP_2D(UMAP_data_7) + #ggtitle("(b)") +
  geom_point(color = "#8dd3c7") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

#(n-neighbors: 50)
plot_list2_umap <- plot_UMAP_2D(UMAP_data_15) + #ggtitle("(b)") + 
  geom_point(color = "#a65628") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

#(n-neighbors: 50)
plot_list3_umap <- plot_UMAP_2D(UMAP_data_32) + #ggtitle("(b)") +
  geom_point(color = "#f781bf") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

#(n-neighbors: 50)
plot_list4_umap <- plot_UMAP_2D(UMAP_data) + #ggtitle("(b)") +
  geom_point(color = "#999999") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'd', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

<!--lwd smoothing criteria with s-curve-->

@@

```{r}
#| echo: false
#| warning: false
#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons. What is the effective benchark valuw to remove the low-density heaxgons?
#| label: fig-diagnosticpltScurvelwd
#| out-width: 100%

@@

@@

>>>>

- Compute high-D distances (using averaged high-D data)
- Fit a linear model using response as distance_high_D and predictor as distance_2D
- Filter the model residuals only for short edges
- compute error for each edge connection as (residual for that point)/(total residual for all edges * number of short edges)

```{r}
#| echo: false
#| warning: false

error_df_lg_1 <- read_rds("data/s_curve/s_curve_summary_lg_tsne.rds")
error_df_lg_2 <- read_rds("data/s_curve/s_curve_summary_lg_umap.rds")
error_df_lg_3 <- read_rds("data/s_curve/s_curve_summary_lg_phate.rds")
error_df_lg_4 <- read_rds("data/s_curve/s_curve_summary_lg_trimap.rds")
error_df_lg_5 <- read_rds("data/s_curve/s_curve_summary_lg_pacmap.rds")

error_df_lg <- dplyr::bind_rows(error_df_lg_1, error_df_lg_2, error_df_lg_3, error_df_lg_4, error_df_lg_5)

error_df_lg$method <- factor(error_df_lg$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))

error_plot_lg_1 <- ggplot(error_df_lg_1, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#e41a1c")) +
  ylab("error") +
  xlab("Benchmark value")

error_plot_lg_2 <- ggplot(error_df_lg_2, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#377eb8")) +
  ylab("error") +
  xlab("Benchmark value")

error_plot_lg_3 <- ggplot(error_df_lg_3, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#4daf4a")) +
  ylab("error") +
  xlab("Benchmark value")

error_plot_lg_4 <- ggplot(error_df_lg_4, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#984ea3")) +
  ylab("error") +
  xlab("Benchmark value")

error_plot_lg_5 <- ggplot(error_df_lg_5, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#ff7f00")) +
  ylab("error") +
  xlab("Benchmark value")
```

```{r}
#| echo: false
#| warning: false
#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the long edges. What is the effective benchmark value to remove the long edges?
#| label: fig-diagnosticpltScurvelgrm
#| out-width: 100%

error_plot_lg_1 + error_plot_lg_2 + error_plot_lg_3 + error_plot_lg_4 + error_plot_lg_5 +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', nrow = 2) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8)) 
```

@@

#(n-neighbors: 50)
plot_list1_umap <- plot_UMAP_2D(UMAP_data_7) + #ggtitle("(b)") +
  geom_point(alpha=0.5, size = 0.5, colour = "#8dd3c7", fill = "#8dd3c7") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

#(n-neighbors: 50)
plot_list2_umap <- plot_UMAP_2D(UMAP_data_15) + #ggtitle("(b)") + 
  geom_point(alpha=0.5, size = 0.5, colour = "#a65628", fill = "#a65628") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

#(n-neighbors: 50)
plot_list3_umap <- plot_UMAP_2D(UMAP_data_32) + #ggtitle("(b)") +
  geom_point(alpha=0.5, size = 0.5, color = "#f781bf", fill = "#f781bf") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

#(n-neighbors: 50)
plot_list4_umap <- plot_UMAP_2D(UMAP_data) + #ggtitle("(b)") +
  geom_point(alpha=0.5, size = 0.5, colour = "#999999", fill = "#999999") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'd', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

<!--lwd smoothing criteria 1 with s-curve-->

@@

```{r}
#| echo: false
#| warning: false
#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons. What is the effective benchmark value to remove the low-density heaxgons?
#| label: fig-diagnosticpltScurvelwd
#| out-width: 100%

@@

https://satijalab.org/seurat/articles/pbmc3k_tutorial.html
https://www.cell.com/fulltext/S0092-8674(15)00549-8

@@

### Retina dataset

https://www.cell.com/cms/10.1016/j.cell.2015.05.002/attachment/fd52950c-fd35-401a-9e60-ca7588ef9001/mmc1.pdf

https://www.cell.com/fulltext/S0092-8674(15)00549-8#app3

32 PCs
tSNE (perplexity 30)

### 2024-01-15 paper.tex
Add logic to remove long edges and add error calculation

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

@@

\caption{\label{fig-diagnosticpltScurvelwd}Goodness of fit statistics
from different NLDR techniques applied to training S-curve dataset with
different benchmark values to remove the low-density heaxgons. What is
the effective benchark valuw to remove the low-density heaxgons?}

@@

@@

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

\begin{itemize}
\tightlist
\item
  Compute high-D distances (using averaged high-D data)
\item
  Fit a linear model using response as distance\_high\_D and predictor
  as distance\_2D
\item
  Filter the model residuals only for short edges
\item
  compute error for each edge connection as (residual for that
  point)/(total residual for all edges * number of short edges)
\end{itemize}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurvelgrm-1.pdf}

\caption{\label{fig-diagnosticpltScurvelgrm}Goodness of fit statistics
from different NLDR techniques applied to training S-curve dataset with
different benchmark values to remove the long edges. What is the
effective benchmark value to remove the long edges?}

@@

\caption{\label{fig-diagnosticpltScurvelwd}Goodness of fit statistics
from different NLDR techniques applied to training S-curve dataset with
different benchmark values to remove the low-density heaxgons. What is
the effective benchmark value to remove the low-density heaxgons?}

@@

https://satijalab.org/seurat/articles/pbmc3k\_tutorial.html
https://www.cell.com/fulltext/S0092-8674(15)00549-8

@@

\hypertarget{retina-dataset}{%
\subsection{Retina dataset}\label{retina-dataset}}

https://www.cell.com/cms/10.1016/j.cell.2015.05.002/attachment/fd52950c-fd35-401a-9e60-ca7588ef9001/mmc1.pdf

https://www.cell.com/fulltext/S0092-8674(15)00549-8\#app3

32 PCs tSNE (perplexity 30)
